\section*{Introduction}
Welcome to the practicals for EEE3096S. These instruction are applicable to all practicals so please take note.

It is critical to do the pre-practical work. Tutors will not help you with questions with answers that would have been known had you done the pre-practical work. 

The practical handbook introduces the important tools, knowledge and skills you will need for these practicals. While references to more detailed information is given, students are encouraged to learn more about these topics using the internet and to take responsibility for their learning.

A special thanks to the tutors and other members who may have contributed to this practical manual by implementing practicals, writing up instructions and generating marking resources. 


\subsection*{Practical Instructions}
Failure to meet these instructions will result in negative marks.
\begin{itemize}
    \item Please check Vula for your practical time. Many practicals require demonstrations, so you will need to attend.
    \item All practicals need to be uploaded to github
    \item All files submitted to Vula need to be in the format
    \begin{lstlisting}[gobble=4]
    pracnum_studnum1_studnum2.fileformat
    \end{lstlisting}
    \item All text assignments must be submitted as pdf
    \item Each pdf should contain a link to your github repository
    \item All \textit{reports} (that is, submissions for Prac 2, Project A and Project B) need to be written in Latex.
    \item Practicals 0 and 1 are to completed alone. Practical 2 onwards is to be completed in groups of two. You can choose your own groups as long as the person is in the same lab session as you. Ideally, your partner should stay the same for the duration of the course.
\end{itemize}